\documentclass{article}
\usepackage[margin=0.7in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsfonts,amsthm}
\usepackage{mathtools}
\usepackage{algorithm}
\usepackage{algpseudocode}

\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}

\newcommand{\pluseq}{\mathrel{+}=}
\newcommand{\minuseq}{\mathrel{-}=}

\newenvironment{claim}[1]{\par\noindent\underline{Claim:}\space#1}{}
\newenvironment{claimproof}[1]{\par\noindent\underline{Proof:}\space#1}{\hfill $\blacksquare$}

\setlength{\parindent}{0pt}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\title{Intelligent Learning and Analysis Systems SS19, Exercise Sheet 2}
\author{by Andreas Hene, Niklas Mertens, Richard Palme.\\
Tutor: Maximilian Thiessen, Group 3}
\date{\today}
\maketitle

\section*{Exercise 3}
\subsection*{Lemma 1}
Each bucket has exactly size $ w = \ceil*{\frac{1}{\varepsilon}}$.
$\Rightarrow$ total number of buckets is
\[
\frac{m}{w} = \frac{m}{\ceil*{1/\varepsilon}} \leq \varepsilon m
\]
So the current bucket id is at most $\varepsilon m$.

\subsection*{Lemma 2}
Proof by induction. \\
Base Case: Let $b_{\text{current}} = 1$, let $(e, f, \Delta) \in \mathcal{D}$.
Since there have been no deletions and no decrementations of $f$ until now,
we know that $f$ is the true frequency up until this point, i.e. $f = f_e$. \\

Also, $(e, f, \Delta)$ is only deleted when $f \leq 1$, in other words $(e, f, \Delta)$ is only deleted when $f_e = f \leq 1 = b_{\text{current}}$ which proves the base case. \\

Step Case: Let $k > 1$. Suppose that for $b_{\text{current}} < k$ we know that $f_e \leq b_{\text{current}}$ when $(e, f, \Delta)$ gets deleted. \\

Now let $b_{\text{current}} = k$ and $(e, f, \Delta) \in \mathcal{D}$.
The true frequency of $e$ in the buckets with ids $\Delta+1, \hdots, b_{\text{current}}$
is $\leq f + (b_{\text{current}}-1 - \Delta)$, because $f$ gets decremented by 1 at most $b_{\text{current}} -1 -\Delta$ times. \\
Let $b'$ be the bucket id where $e$ was deleted the last time, if it exists.
Else set $b'=0$. By the induction hypothesis, the true frequency of $e$ in buckets
$1, \hdots, b'$ is $\leq b'$, if $b'$ exists. The true frequency in buckets
$b'+1, \hdots, \Delta$ is $0$. So the true frequency $f_e$ at the current time
(bucket id $b_{\text{current}}$) is at most $f_e \leq f + b_{\text{current}}-1 - \Delta + b'$. \\
Since $b' \leq \Delta$, we get: $f_e \leq f + b_{\text{current}} - 1$.\\

If $(e, f, \Delta)$ gets deleted, we have $f \leq 1$, so $f_e \leq f+b_{\text{current}}-1 \leq b_{\text{current}}$.

\subsection*{Lemma 3}
If there is no entry for $e$ in $\mathcal{D}$, then there are 2 cases:\\
Case 1: There has never been an entry for $e$ in $\mathcal{D}$. \\
Then $f_e = 0 \leq \varepsilon m$.\\
Case 2: There has been an entry for $e$ in $\mathcal{D}$ before.\\
Let $b$ be the bucket id when $e$ was deleted the last time. Then $f_e \leq b$ by Lemma 2, and $f_e \leq b \leq \varepsilon m $ by Lemma 1.

\subsection*{Lemma 4}
$f \leq f_e$, because $f$ gets incremented by 1 at most $f_e$ times. \\
If the last time $e$ was deleted was after processing bucket $b$, then by Lemma 3 the true frequency of $e$ in buckets
$1, \hdots, b$ is at most $\varepsilon m$. The true frequency of $e$ in buckets
$b+1, \hdots, \Delta$ is 0, and the true frequency in buckets
$\Delta+1, \hdots, b_{\text{current}}$ is  at most $f$. So:
\[
f_e \leq \varepsilon m + 0 + f = f + \varepsilon m
\]

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Exercise 4}
$d_i$ is the number of elements of $\mathcal{D}$ that were last added to $\mathcal{D}$ during processing of bucket $B-i+1$. Call the i-th summand the contribution of bucket $B-i+1$ to the sum. We want that the contribution of all summands does not exceed the total size of buckets $B-j+1, \hdots, B$ (which is $jw$). \\

So if element $e$ was created during processing of $B-i+1$, then $e$ survives the
deletion/decrementing process $i-1$ times. This can only happen if $e$ occurs at least $i$ times in buckets $B-i+1, \hdots B$. So $e$ is allowed a contribution of $i$. So the contribution of all the elements last added to $\mathcal{D}$ during processing of bucket $B-i+1$ is allowed to be $id_i$, because this way the contribution of all summands can't exceed the total size of $B-j+1, \hdots B$, i.e.
\[
\sum_{i=1}^B id_i \leq jw
\]

\newpage
\section*{Exercise 2}
Let $m_1 \coloneqq \lvert \sigma_1 \rvert ,m_2 \coloneqq \lvert \sigma_2 \rvert, m \coloneqq m_1 + m_2$\\
$m'_1 \coloneqq \sum_{l=1}^{k-1}A_1[l], m'_2 \coloneqq \sum_{l=1}^{k-1}A_2[l], m' \coloneqq \sum_{l=1}^{k-1}A[l]$.\\
$c_k \coloneqq $ counter of k-th most frequent item in $A_1 \oplus A_2$\\
Let $i \in \text{keys}(A).$\\
The estimated frequency of $i$ in $\sigma_1$ is at most $\frac{m_1 - m'_1}{k}$ smaller than the true frequency of $i$ in $\sigma_1$.\\
The estimated frequency of $i$ in $\sigma_2$ is at most $\frac{m_2 - m'_2}{k}$ smaller than the true frequency of $i$ in $\sigma_2$.\\
This the estimated frequency of $i$ in $\sigma_1 \otimes \sigma_2$ is at most $\frac{m_1 - m'_1}{k} + \frac{m_2 - m'_2}{k} + c_k$ smaller than the true frequency of $i$ in $\sigma_1 \otimes \sigma_2$.\\
\[
\Rightarrow \hat{f}_i \geq f_i - \frac{m-(m'_1 + m'_2 -kc_k)}{k}
\]
\begin{claim}
$m'_1 + m'_2 -kc_k \geq m'$
\end{claim}
\begin{claimproof}
Case 1: $\lvert \text{keys}(A_1) \cup \text{keys}(A_2) \rvert \leq k-1$\\
Then $c_k = 0$ and $m' = m'_1 + m'_2$.

Case 2: $\lvert \text{keys}(A_1) \cup \text{keys}(A_2) \rvert \geq k$\\
In line 5 of Misra-Gries synopses algorithm we have $m'_1 + m'_2 \geq \widetilde{m}' + c_k$,
since the k-th most frequent item is not in $A$ ($\widetilde{m}'$ denotes $m'$ at this point of the algorithm).\\
In line 7 of the algorithm, $\widetilde{m}'$ is reduced by $(k-1)c_k$, i.e.
\[
m' = \widetilde{m}' - (k-1)c_k \Rightarrow m'_1 + m'_2 \geq m' + kc_k
\]
\end{claimproof}

So
\[
\hat{f}_i \geq f_i - \frac{m-m'}{k}
\]

$\hat{f}_i \leq f_i$, because $A_1[i]$ and $A_2[i]$ are underestimations of the true frequencies of $i$ in $\sigma_1, \sigma_2$.
So $A_1[i] + A_2[i]$ is an underestimation of $f_i$ in $\sigma_1 \otimes \sigma_2$, so $A[i]$ is an underestimation of $f_i$ as well.


\end{document}
